Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-09-13 ms.tex
Note for elaboration between computational methods and observational background
Introducing numerical simulations paragraph better, edits through that paragraph of the intro

Hubble Space Telescope (HST) observations with STIS have found large \lya transit depths around a handful of exoplanets \citep{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order a few planetary radii for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with the H atom population. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in the atmospheric dynamics of these systems as the optical depth at line center is large. The importance of the \lya line in this process necessitates a careful treatment of resonant scattering in order to construct accurate models of H atom excitation, heating, and radiative forces. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods.

The main impediment to performing magnetohydrodynamic simulations with fully a coupled radiative transfer solution via Monte Carlo methods is the prohibitive computational cost. A primary factor in the cost of Monte Carlo is that the number of scatterings a photon undergoes is proportional to the line center optical depth.  Near the base of the atmosphere, individual simulation cells can have optical depths of more than a million, so most of the time is spent integrating photons in these cells. This means that any method that can treat these zone via an analytic solution has the potential to greatly accelerate the calculation. A complete, general analytic solution for the resonance scattering has never been derived but approximate solutions exist in certain limits.

>>>>

Hubble Space Telescope (HST) observations with the STIS have found large \lya transit depths around a handful of exoplanets \citep{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order a few planetary radii for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with the H atom population. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in the atmospheric dynamics of these systems as the optical depth at line center is large. The importance of the \lya line in this process necessitates a careful treatment of resonant scattering in order to construct accurate models of H atom excitation, heating, and radiative forces.

Numerical simulations can be used to understand the complex layers of these irradiated exoplanet atmospheres. However, the extreme optical depths at \lya line center introduce a computational barrier to simulating radiative transfer via Monte Carlo methods. The main impediment to performing magnetohydrodynamic simulations with fully a coupled radiative transfer solution via Monte Carlo methods is the prohibitive computational cost. The number of scatterings a photon undergoes is proportional to the line center optical depth.  Near the base of the atmosphere, the line center optical depth is ${\sim}10^9$, so most of the time is spent integrating photons in these cells. A method that can treat these zones via a simple solution has the potential to greatly accelerate the calculation. A complete, general analytic solution for the resonance scattering has never been derived but approximate solutions exist in certain limits.
